Efficient pedigree recording for fast population genetics simulation

### 2018-01-12 forwards_paper.tex
Minor updates (+ remove trailing whitespace).

\begin{abstract}
    In this paper we describe how to
    efficiently record the entire genetic history of a population
    during a forwards-time, individual-based population genetics simulation.
    This dramatically reduces the computational burden of tracking individual genomes
    by simulating only those loci that may affect reproduction (those having non-neutral variants),
    and recording history as a \emph{succinct tree sequence} as introduced in the software package \msprime,
    on which neutral mutations can be quickly placed afterwards.
    We implement the method in two popular forwards-time simulation frameworks,
    which speeds up large, whole-genome simulations by one to orders of magnitude.
    In addition to speed, the method has several advantages:
    (1) All marginal genealogies of the simulated individuals are recorded, rather than just genotypes.
    (2) A population of $N$ individuals with $M$ polymorphic sites
    can be stored in $O(N \log N + M)$ space, making it feasible to store a simulation's entire final generation (and its' history)
    rather than only a subset.
    (3) A simulation can easily be initialized with a more efficient coalescent simulation of deep history.
    To make these possible,
    we introduce the \msprime{} Tables API, which allows for the efficient interchange 
    of tree sequences between modular program components,
    and show how to quickly simplify a tree sequence by removing history irrelevant to a given set of genomes.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
Thanks to Jared Galloway, Brad Shaffer, and Evan McCartney--Melstad for useful discussions.
Work on this project was supported by funding from 
the Sloan Foundation and the NSF (under DBI-1262645) to PR;
and XXX.

>>>>

\begin{abstract}
    In this paper we describe how to
    efficiently record the entire genetic history of a population
    during a forwards-time, individual-based population genetics simulation.
    This dramatically reduces the computational burden of tracking individual genomes
    by simulating only those loci that may affect reproduction (those having non-neutral variants),
    and recording history as a \emph{succinct tree sequence} as introduced in the software package \msprime,
    on which neutral mutations can be quickly placed afterwards.
    We implement the method in two popular forwards-time simulation frameworks,
    which speeds up large, whole-genome simulations by one to two orders of magnitude.
    In addition to speed, the method has several advantages:
    (1) All marginal genealogies of the simulated individuals are recorded, rather than just genotypes.
    (2) A population of $N$ individuals with $M$ polymorphic sites
    can be stored in $O(N \log N + M)$ space, making it feasible to store a simulation's entire final generation (and its' history)
    rather than only a subset.
    (3) A simulation can easily be initialized with a more efficient coalescent simulation of deep history.
    To make these possible,
    we introduce the \msprime{} Tables API, which allows for the efficient interchange
    of tree sequences between modular program components,
    and show how to quickly simplify a tree sequence by removing history irrelevant to a given set of genomes.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
Thanks to Jared Galloway, Brad Shaffer, and Evan McCartney--Melstad for useful discussions.
Work on this project was supported by funding from
the Sloan Foundation and the NSF (under DBI-1262645) to PR;
the Wellcome Trust (grant 100956/Z/13/Z to Gil McVean);
and XXX.
